---
title: "Module 1 Homework"
author: "Filipp Krasovsky"
date: "3/8/2021"
output:
  pdf_document: default
  pdf: default
---

Assignment:
Data Science Using Python and R: Chapter 3 - Page 45: Questions #21, 22, 23, 24, & 25 
Hint: Use both R and Python for these questions.
Hint: Datasets for this assignment are available on the Weekly Python & R with Datasets page or you can download them here.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



For Exercises 21–25, work with the Nutrition_subset data set. The data set contains the
weight in grams along with the amount of saturated fat and the amount of cholesterol for a
set of 961 foods. Use either Python or R to solve each problem.



<b>Question 21 - Ch.3</b>
The elements in the data set are food items of various sizes, ranging from a teaspoon of
cinnamon to an entire carrot cake.
a. Sort the data set by the saturated fat (saturated_ fat) and produce a listing of the five
food items highest in saturated fat.
b. Comment on the validity of comparing

```{r}
#Chapter 3

#read our nutritional data in 
df = read.csv("C:/Users/Filipp/Documents/usd_data_sci/502_data mining/module1/Website Data Sets/nutrition_subset.csv")
```

```{r}
#QUESTION 21 

#a:sort by saturated fat 
df.sorted_desc = df[order(df$saturated_fat,decreasing=TRUE),]

#get top five food items 
df.top_five = head(df.sorted_desc,n=5)

#print our top five
print(df.top_five)
```

b. the comparison of saturated fat is not meaningful because we are looking at different serving sizes, even across the same food group (ex: Parmesan Cheese).

<b>Question 22 - Ch. 3 </b>
Derive a new variable, saturated_ fat_per_gram, by dividing the amount of saturated fat
by the weight in grams.
a. Sort the data set by saturated_ fat_per_gram and produce a listing of the five food
items highest in saturated fat per gram.
b. Which food has the most saturated fat per gram?

```{r}
#create our new variable 
df$saturated_fat_per_gram = df$saturated_fat/df$weight_in_grams
#a. sort by saturated fat per gram 
df.sorted_by_sfpg = df[order(df$saturated_fat_per_gram,decreasing=TRUE),]
#b. get the food with the most saturated fat per gram 
most_sfpg = subset(head(df.sorted_by_sfpg,n=1),select=c(food.item,saturated_fat_per_gram))

print(most_sfpg)
```
<b> Question 23 - Ch. 3 </b>

Derive a new variable, cholesterol_per_gram.
a. Sort the data set by cholesterol_per_gram and produce a listing of the five food
items highest in cholesterol fat per gram.
b. Which food has the most cholesterol fat per gram?

```{r}
#create new variable 
df$cholesterol_per_gram = df$cholesterol/df$weight_in_grams

#sort dataset, produce top five, output food with the most cholesterol 
df.sorted_chol = df[order(df$cholesterol_per_gram,decreasing=TRUE),]
df.top_five_chol = subset(head(df.sorted_chol,n=5),select=c(food.item,cholesterol_per_gram))

print(df.top_five_chol)

```
The food item with the most cholesterol per gram is Eggs.

<b> Question 24 - ch.3 </b>

Standardize the field saturated_fat_per_gram. Produce a listing of all the food items
that are outliers at the high end of the scale. How many food items are outliers at the low
end of the scale?

```{r}
#begin by looking at distribution
d <- density(df$saturated_fat_per_gram) # returns the density data
plot(d)
```
Our plot suggests that saturated_fat_per_gram does not follow a standard normal distribution, but rather a chi-square distribution of k~3. There are no outliers at the low end of the scale. Based on the proceeding chi-square test for outliers, we can conclude that we have at least one outlier on the high end. There are no outliers on the low end.
```{r}
require(outliers)
chisq.out.test(df$saturated_fat_per_gram)

#standardize 
df$sfpg.std = scale(x = df$saturated_fat_per_gram)
subset(df,abs(df$sfpg.std) >= 3)
```
Question 25 - ch. 3
Standardize the field cholesterol_per_gram. Produce a listing of all the food items that
are outliers at the high end of the scale.
```{r}
df$cpg_z <- scale(df$cholesterol_per_gram)
subset(df,(df$cpg_z) >= 3)
```


