---
title: "Module 5 Exercise 2"
author: "Filipp Krasovsky"
date: "11/30/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

2.2 Use R to understand how horsepower and weights are related to each other. Plot
them using a scatter plot and color the data points using mpg. Do you see anything
interesting/useful here? Report your observations with this plot. Now let us cluster the data on this plane in a “reasonable” number of groups. Show your plot where the data points are now colored with the cluster information and provide your interpretations. (10 points)

```{r}
#install packages
library(RMySQL)

#connect to db
default_authentication_plugin='admin'
db = dbConnect(MySQL(),user='root',password='admin', dbname='500b', host='127.0.0.1')

#extract mpg data.
mpg_df = fetch(dbSendQuery(db,"SELECT * FROM mpg"),n=-1)
mpg_df = as.data.frame(mpg_df)

#print head for sanity check
head(mpg_df)
```

```{r}
#plot horsepower and weight, color with mpg.
library(ggplot2)
ggplot(mpg_df,aes(horsepower,weight,color=mpg)) + geom_point()
```
Initial observations demonstrate two major points of interest:
1. there is a very strong positive correlation between weight and horsepower.
This is rather intuitive as the heavier a vehicle is, the more horsepower is required to propel it.

2. Combinations of horsepower and weight with increased values for both are associated, on face value, with a lower mpg. 

Given that the gradient of mpg falls between ~10 and ~50, we can create 4 different clusters for our kmeans.

```{r}
set.seed(20)
mpgCluster <- kmeans(mpg_df[,4:5],4,nstart=20)

ggplot(mpg_df,aes(horsepower,weight,color=mpgCluster$cluster)) + geom_point()
```

In this instance, the output of the kmeans clustering process can be interpreted as follows:
Cluster 1 = 20-30 mpg
Cluster 2 = 10-20 mpg
Cluster 3 = 40-50 mpg
Cluster 4 = 30-40 mpg

Overall, the learning model captures the noise in the data well but makes
some mistakes at smaller values of horsepower/weight.


